---
title: "U.S. Region Tuition Increase Comparison"
author: "Jayne Hoyt"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    code_folding: show
    number_sections: true
    theme: spacelab
    highlight: tango
    fig_caption: true
    keep_md: true
    self_contained: false
---
```{r, include=FALSE}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(caret)
library(beepr)
library(readxl)
library(tseries)
library(forecast)
library(haven)
library(fma)
library(expsmooth)
library(lmtest)
library(zoo)
library(aTSA)
library(imputeTS)
library(reticulate)
library(lubridate)
```

# Loading in Data

```{r}

newEngland <- read.csv("4YrPublicNewEngland.csv")
newEngland <- newEngland %>%
  filter(Year >= 2002)

national <- read.csv("4YrPublicNational.csv")
national <- national %>%
  filter(Year >= 2002)

middle <- read.csv("4YrPublicMiddle.csv")
middle <- middle %>%
  filter(Year >= 2002)

midwest <- read.csv("4YrPublicMidwest.csv")
midwest <- midwest %>%
  filter(Year >= 2002)

west <- read.csv("4YrPublicWest.csv")
west <- west %>%
  filter(Year >= 2002)

south <- read.csv("4YrPublicSouth.csv")
south <- south %>%
  filter(Year >= 2002)

southwest <- read.csv("4YrPublicSouthwest.csv")
southwest <- southwest %>%
  filter(Year >= 2002)

fourYrPublic <- read.csv("4YrPublicALL.csv")
fourYrPublic <- fourYrPublic %>%
  filter(Year >= 2002)


head(fourYrPublic)


```
# Vizualizing Data

```{r}
#Tuition Prices over Time Among Regions
plot <- ggplot(fourYrPublic, aes(Year)) +
  geom_line(aes(y = MiddleStates), color = "#e40000", linetype = "solid", linewidth = 1.5) +
  geom_line(aes(y = Midwest), color = "#FA9C1B", linetype = "solid", linewidth = 1.5) +
  geom_line(aes(y = NewEngland), color = "#FFD700", linetype = "solid", linewidth = 1.5) +
  geom_line(aes(y = South), color = "#90EE90", linetype = "solid", linewidth = 1.5) +
  geom_line(aes(y = Southwest), color = "#ADDAE6", linetype = "solid", linewidth = 1.5) +
  geom_line(aes(y = West), color = "#D6B4Fc", linetype = "solid", linewidth = 1.5) +
  labs(x = "Year", y = "Tuition Cost", title = "Tuition Prices over Time Among Regions (2002-2023)") +
  theme_dark() + theme(
  text = element_text(family = "sans", size = 12),  # Change font to Arial and set size
  title = element_text(family = "serif", face = "bold", size = 16),  # Change title font
  axis.text.x = element_text(family = "sans", angle = 45, hjust = 1),  # Change x-axis text font and rotation
  axis.text.y = element_text(family = "sans", size = 10),  # Change y-axis text font
  legend.text = element_text(family = "serif", size = 8)  # Change legend text font
)

plot
```



------------------------------------

# Linear Regression for Each Region

## National
```{r}
national_lm  <- lm( Cost ~ Year, data = national)
summary(national_lm)
```

## Middle States
```{r}
middle_lm  <- lm(Cost ~ Year, data = middle)
summary(middle_lm)
```

## Midwest
```{r}
midwest_lm  <- lm(Cost ~ Year, data = midwest)
summary(midwest_lm)
```



## New England
```{r}
newEngland_lm  <- lm(Cost ~ Year, data = newEngland)
summary(newEngland_lm)
```


## South
```{r}
south_lm  <- lm(Cost ~ Year, data = south)
summary(south_lm)
```


## Southwest

```{r}
southwest_lm  <- lm(Cost ~ Year, data = southwest)
summary(southwest_lm)
```

## West
```{r}
west_lm  <- lm(Cost ~ Year, data = west)
summary(west_lm)
```





# Performing Two-Sample T-Tests


## Southwest v West

```{r}
#Southwest v West
# Null : Southwest = West
# Alternative: Southwest < West 

 #(beta1_{group1} - beta1_{group2})/sqrt(se(beta1__{group1})^2 + se(beta1_{group2})^2)

t = (344.9 - 375.97) / sqrt((9.272)^2 + (21.01)^2)
t
```
Test Statistic = -1.353

Degrees of Freedom= 20

```{r}
#Find P-value
p = 2*pt(t, 20) 
p = p * 15 # Bonferonni adjustment
p
```
P-value : 1

*Not significantly different*

## **West v New England**

```{r}
#West vs New England
# Null : West = New England
# Alternative: West < New England

t = (375.97- 466.6) / sqrt((21.01)^2 + (7.986)^2)
t
```
Test Statistic = -4.03

Degrees of Freedom= 20

```{r}
#Find P-value
p = 2*pt(-4.03, 20) 
p = p * 15  #Bonferroni adjustment 
p

```
P-value = 0.00984

**SIGNIFICANTLY DIFFERENT**

Therefore we reject the null hypothesis. There is sufficient evidence to support the claim that the West is increasing a statistically lower rate than New England.


## Midwest v West

```{r}
# Midwest vs West
# Null :Midwest = West 
# Alternative: Midwest < West 

t = (322.56 - 375.97) / sqrt((10.94)^2 + (21.01)^2)
t
```
Test Statistic = -2.255

Degrees of Freedom = 20

```{r}
#Find P-value
p = 2*pt(t, 20)
p = p * 15  #Bonferroni adjustment 
p
```
P-value = 0.532

*Not significantly different* 

## **Southwest v New England**

```{r}
#Southwest vs New England
# Null : Southwest = New England
# Alternative: Southwest < New England

t = (344.9- 466.6) / sqrt((9.272)^2 + (7.986)^2)
t
```
Test Statistic = -9.945

Degrees of Freedom = 20

```{r}
#Find P-value
p = 2*pt(t, 20) 
p = p * 15 # Bonferonni adjustment
p
```
P-value = 5.205e-08

**SIGNIFICANTLY DIFFERENT**

Therefore we reject the null hypothesis. There is sufficient evidence to support the claim that the Southwest is increasing a statistically lower rate than New England.


##  Midwest v Southwest
```{r}
#Midwest vs Southwest
# Null : Midwest = Southwest
# Alternative: Midwest < Southwest

t = (322.56 - 344.9) / sqrt((10.94)^2 + (9.272)^2)
t
```
Test Statistic = -1.56

Degrees of Freedom = 20

```{r}
#Find P-value
p = 2*pt(t, 20) 
p = p * 15 # Bonferonni adjustment
p
```
P-value = 1

*Not significantly different*


## South v West

```{r}
#South v West
# Null : South = West
# Alternative: South < West

t = (344.83 - 375.97) / sqrt((15.44)^2 + (21.01)^2)
t
```
Test Statistic = -1.194

Degrees of Freedom = 20

```{r}
#Find P-value
p = 2*pt(t, 20) 
p = p * 15 # Bonferonni adjustment
p
```
P-value = 1

*Not significantly different* 

## South v Southwest
```{r}
#South vs Southwest
# Null : South = Southwest
# Alternative: South < Southwest

t = (344.83 - 344.9) / sqrt((9.272)^2 + (15.44)^2)
t
```
Test Statistic = -0.00389

Degrees of Freedom = 20

```{r}
#Find P-value
p = 2*pt(t, 20) 
p = p * 15 # Bonferonni adjustment
p 
```
P-value = 1

*Not significantly different*


## **South v New England**
```{r}
#South vs New England
# Null : South = New England
# Alternative: South < New England

t = (344.83- 466.6) / sqrt((15.44)^2 + (7.986)^2)
t
```
Test Statistic = -7.005

Degrees of Freedom = 20

```{r}
#Find P-value
p = 2*pt(t, 20) 
p = p * 15 # Bonferonni adjustment
p 

```
P-value = 1.276e-05

**SIGNIFICANTLY DIFFERENT**

Therefore we reject the null hypothesis. There is sufficient evidence to support the claim that the South is increasing a statistically lower rate than New England.

## Midwest v South
```{r}
#Midwest vs South
# Null :  Midwest = South
# Alternative: Midwest < South

t = (322.56 - 344.83) / sqrt((10.94)^2 + (15.44)^2)
t
```
Test Statistic = -1.177

Degrees of Freedom = 20

```{r}
#Find P-value
p = 2*pt(t, 20) 
p = p * 15 # Bonferonni adjustment
p
```
P-value = 1

*Not significantly different*


## South v Middle States
```{r}
#Middle States vs South
# Null :Middle States  = South
# Alternative:Middle States  < South

t = (319.3 - 344.83) / sqrt((6.675)^2 + (15.44)^2)
t
```
Test Statistic = -1.518

Degrees of Freedom = 20

```{r}
#Find P-value
p = 2*pt(t, 20) 
p = p * 15 # Bonferonni adjustment
p
```
P-value = 1

*Not significantly different* 


## **Midwest v New England**
```{r}
# Midwest vs New England
# Null : Midwest = New England
# Alternative: Midwest < New England

t = (322.56 - 466.6) / sqrt((7.986)^2 + (10.94)^2)
t
```
Test Statistic = -10.634

Degrees of Freedom = 20

```{r}
#Find P-value
p = 2*pt(t, 20) 
p = p * 15 # Bonferonni adjustment
p
```
P-value = 1.671e-08

**SIGNIFICANTLY DIFFERENT**

Therefore we reject the null hypothesis. There is sufficient evidence to support the claim that the Midwest is increasing a statistically lower rate than New England.

## Middle States v West
```{r}
#Middle States v West
# Null : Middle States = West
# Alternative: Middle States < West

t = (319.3 - 375.97) / sqrt((21.01)^2 + (6.675)^2)
t
```
Test Statistic = -2.57

Degrees of Freedom = 20

```{r}
#Find P-value
p = 2*pt(t, 20) 
p = p * 15 # Bonferonni adjustment
p
```
P-value = 0.2737

*Not significantly different* 

## Middle States v Southwest

```{r}
#Middle States v Southwest
# Null : Middle States = Southwest
# Alternative: Middle States < Southwest
 #(beta1_{group1} - beta1_{group2})/sqrt(se(beta1__{group1})^2 + se(beta1_{group2})^2)

t = (319.3 - 344.9) / sqrt((9.272)^2 + (6.675)^2)
t
```
Test Statistic = -2.241

Degrees of Freedom = 20

```{r}
#Find P-value
p = 2*pt(t, 20) 
p = p * 15 # Bonferonni adjustment
p
```
P-value = 0.548

*Not significantly different*

## **Middle States v New England**
```{r}
#Middle States v New England
# Null : Middle States = New England
# Alternative: Middle States < New England
 #(beta1_{group1} - beta1_{group2})/sqrt(se(beta1__{group1})^2 + se(beta1_{group2})^2)

(319.3 - 466.6) / sqrt((7.986)^2 + (6.675)^2)
```
Test Statistic = -7.387

Degrees of Freedom = 20

```{r}
#Find P-value
2*pt(-7.387, 20)
```
P-value = 1.745E-08

**SIGNIFICANTLY DIFFERENT**

Therefore we reject the null hypothesis. There is sufficient evidence to support the claim that the Middle States are increasing a statistically lower rate than New England.

## Middle States v Midwest
```{r}
#Middle States v Midwest
# Null : Middle States = Midwest
# Alternative: Middle States < Midwest
 #(beta1_{group1} - beta1_{group2})/sqrt(se(beta1__{group1})^2 + se(beta1_{group2})^2)

(319.3 - 322.56) / sqrt((10.94)^2 + (6.675)^2)
```
Test Statistic = -2.007

Degrees of Freedom = 20

```{r}
#Find P-value
2*pt(-2.007, 20)
```
P-value = 0.0530

*Not significantly different*




# Findings

After completing this analysis, I found that :


- West vs New England
- Southwest vs New England
- South vs New England
- Midwest vs New England
- Middle States vs New England 

are all significantly different using a p-value of 0.01.


# Future Investigations

- Look into the cause of why New England is rising at a higher rate than other regions.